\section{Easy peasy lemon squeezy}
Suppose that we have collected the data of a group of students of ML course so that this data includes two features. The first characteristic, $x_1$, is equal to the total number of hours that the student has practiced ML. The second characteristic, $x_2$, is the student's total grade point average (GPA) before taking the ML course. Now, we are going to use a logistic regression model to predict the probability that a student will get a score of 20 in this course. After learning based on these data, the obtained coefficients are equal to $\beta_0 = -5$, $\beta_1 = 0.1$ and $\beta_2 = 0.25$.
\subsection{part 1}
Calculate the probability that a student with 80 hours of study and an GPA of 18 can get a score of 20 in this course.
\begin{qsolve}
    \begin{qsolve}[]
        we can model this problem as a binary logistic regression problem. we know that for a binary logistic regression problem we have:
            $$P(y=1|x;\theta) = \sigma (\alpha )$$
        where $\sigma$ is the sigmoid function and $\alpha = w^T x +b$ where $w$ is the weight vector and $b$ is the bias and $x$ is the input vector. now if we say y is the probability of getting a score of 20 in this course we can write:
            $$P(y=1|x;\theta) = \sigma (\beta_0 + \beta_1 x_1 + \beta_2 x_2)$$
            $$x = \begin{bmatrix} 1 \\ x_1 \\ x_2 \end{bmatrix} = \begin{bmatrix} 1 \\ 80 \\ 18 \end{bmatrix}$$
            $$w = \begin{bmatrix} \beta_0 \\ \beta_1 \\ \beta_2 \end{bmatrix} = \begin{bmatrix} -5 \\ 0.1 \\ 0.25 \end{bmatrix}$$
            $$P(y=1|x;\theta) = \sigma (-5 + 0.1 \times 80 + 0.25 \times 18) = \sigma(7.5) = \frac{1}{1+e^{-7.5}} = 0.9994$$
    \end{qsolve}
\end{qsolve}
\subsection{part 2}
Consider another student who has a GPA of 16 and wants to get a grade of 20 in this course. According to the model that we have taught, how many hours should he practice in order to achieve this grade with a 90\% probability?
\begin{qsolve}
    \begin{qsolve}[]
        we use the previous model in this part too. we have:
        $$x = \begin{bmatrix} 1 \\ x_1 \\ x_2 \end{bmatrix} = \begin{bmatrix} 1 \\ x_1 \\ 16 \end{bmatrix}$$
        $$w = \begin{bmatrix} \beta_0 \\ \beta_1 \\ \beta_2 \end{bmatrix} = \begin{bmatrix} -5 \\ 0.1 \\ 0.25 \end{bmatrix}$$
        $$P(y=1|x;\theta) = \sigma(\alpha) \geq 0.9 \Rightarrow \frac{1}{1+e^{-\alpha}} \geq 0.9$$
        $$\Rightarrow e^{-\alpha} \leq \frac{1}{9} \Rightarrow \alpha \geq ln(9) $$
        $$\Rightarrow \beta_0 + \beta_1 x_1 + \beta_2 x_2 \geq ln(9)$$
        $$\Rightarrow -5 + 0.1 x_1 + 0.25 \times 16 \geq ln(9)$$
        $$\Rightarrow 0.1 x_1 \geq ln(9) + 1 \Rightarrow x_1 \geq \frac{ln(9) + 1}{0.1} = 31.97$$
    \end{qsolve}
\end{qsolve}
